---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Showing code that is required to load the data into the directory
```{r, echo = TRUE}
activitydf <- read.csv('~/Downloads/activity.csv', sep = ',', header = TRUE)
```
Good to see if the data has actually loaded
```{r, echo = TRUE}
head(activitydf)
```

Then we take time to view the structure of the loaded data
```{r, echo = TRUE}
str(activitydf)
```
## What is mean total number of steps taken per day?

Make a histogram of the total number of steps taken each day

We need to load the dplyr package to so as to access the group_by function which will help us group the data by date
```{r, echo = TRUE}
library(dplyr)
```

To calculate the total number of steps taken per day, we first group the data by date, then summarise it with the sum function.
```{r, echo = TRUE}
perday <- activitydf %>%
        group_by(date = as.Date(date)) %>%
        summarise(steps = sum(steps, na.rm = TRUE))
```
We can then plot a histogram after grouping the data
```{r, echo = TRUE}
hist(perday$steps, main = "Total Number of Steps Taken Per Day", xlab = "Steps")     
```

The we can calculate and report the mean and median of the total number of steps taken per day

```{r, echo = TRUE}
mean(perday$steps, na.rm = TRUE)
median(perday$steps, na.rm = TRUE)
```


## What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
First we group the data by interval, then summarise it on the steps variable by its mean and save it into a new object
```{r, echo = TRUE}
avgsteps <- activitydf %>%
        group_by(interval) %>%
        summarise(steps = mean(steps, na.rm = TRUE))
```
From here its easy to plot the graph of the result in the new object.
But first we call the plotting graph device ggplot
```{r, echo = TRUE}
library(ggplot2)
```
Then we are able to plot the result using the ggplot graph ploting device
```{r, echo = TRUE}
ggplot(avgsteps, aes(interval, steps)) + geom_line() + 
        labs(title = "Average Daily Activity Pattern" ) +
        xlab("5-Minute Intervals") + 
        ylab("Average Number of steps taken") + 
        theme(panel.grid.minor = element_line(colour = "blue", linetype = "dotted"
```


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
