---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Showing code that is required to load the data into the directory
```{r, echo = TRUE}
activitydf <- read.csv('~/Downloads/activity.csv', sep = ',', header = TRUE)
```
Good to see if the data has actually loaded
```{r, echo = TRUE}
head(activitydf)
```

Then we take time to view the structure of the loaded data
```{r, echo = TRUE}
str(activitydf)
```
## What is mean total number of steps taken per day?

Make a histogram of the total number of steps taken each day

We need to load the dplyr package to so as to access the group_by function which will help us group the data by date
```{r, echo = TRUE}
library(dplyr)
```

To calculate the total number of steps taken per day, we first group the data by date, then summarise it with the sum function.
```{r, echo = TRUE}
perday <- activitydf %>%
        group_by(date = as.Date(date)) %>%
        summarise(steps = sum(steps, na.rm = TRUE))
```
We can then plot a histogram after grouping the data
```{r, echo = TRUE}
hist(perday$steps, main = "Total Number of Steps Taken Per Day", xlab = "Steps")     
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
